---
title: "w241_tweets_facts_sentiment_FN"
author: "Arriaga, Hua, Khural, Moran"
date: "July 31, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(data.table)
library(foreign)
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(gridExtra)
library(stargazer)                      # for printing tables 

theme_set(theme_classic())
# Set working directory
# setwd("C:/Users/Zaptetra/Google Drive/MIDS/W241/w241_final_project/notebooks")
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 7)
```

## The raw data

First we will load the data.
We will process one dataset at a time.

```{r load data}
d<- data.table::fread('../data/raw/response_tp_07_31.csv')
d<- data.table::fread('../data/raw/response_tn_07_31.csv')
d<- data.table::fread('../data/raw/response_fp_07_31.csv')
d<- data.table::fread('../data/raw/response_fn_07_31.csv')
```

## Data cleaning

We will first review:

- Eliminate empty columns
- Rename columns
- Update fields that didn't export correctly (missing data)
- Check for attrition or duplicates
- Check for fake answers: response time too quick plus attention check question wrong.
- Convert data to correct type (all are strings)

```{r pressure, echo=FALSE}

# delete unwanted rows
unwanted_rows <- c(1,2)

# # Drop blank columns
drop_columns <- c("RecipientLastName",
                  "RecipientFirstName",
                  "RecipientEmail",
                  "ExternalReference", "Progress", "Status",
                  "LocationLatitude",
                  "LocationLongitude","DistributionChannel",
                  "UserLanguage","IPAddress")

drop_col_index <- which(colnames(d) %in% drop_columns)

d<- d[-unwanted_rows, ]

d[, drop_columns]<-list(NULL)

col_new_names <- c("start_date", "end_date", "survey_duration",
                  "finished", "finished_date", "id", "email",
                  "gender", "age", "education",
                  "tweet_first_click_time", "tweet_last_click_time", "tweet_submit_time", "tweet_click_count",
                  "math_q1", "math_q2",
                  "test_first_click_time", "test_last_click_time", "test_submit_time", "test_click_count",
                  "stimulus","georgians","energy","soccer","fauci", "pollution","election")

# Rename columns
colnames(d) <- col_new_names

# Check for missing data
# Audit missing values
####################
# Survey id: R_3s64ZlPIAtROugC - missing soccer and fauci (response in qualtrics confirmed)
d[id=="R_3s64ZlPIAtROugC"]$soccer <- "None of the above"
d[id=="R_3s64ZlPIAtROugC"]$fauci <- "None of the above"

# survey id: R_33vRD5yKc2c3FIk - missing energy (response confirmed)
d[id=="R_33vRD5yKc2c3FIk"]$energy <- "None of the above"

# # survey id: R_3j7jocPTJApWZXu - missing fauci
d[id=="R_3j7jocPTJApWZXu"]$fauci <- "None of the above"

# survey id: R_2zoqcnPiyzpmJga - missing fauci
d[id=="R_2zoqcnPiyzpmJga"]$fauci <- "None of the above"

# survey id: R_1I76YKkC3bSzfJv - missing fauci
d[id=="R_1I76YKkC3bSzfJv"]$pollution <- "None of the above"
d[id=="R_1I76YKkC3bSzfJv"]$georgians <- "None of the above"

# survey id: R_2asjDNe8PAuhxxh - missing fauci
d[id=="R_2asjDNe8PAuhxxh"]$georgians <- "None of the above"

# survey id: R_d4o2OrwcVxrjEn7 - missing fauci
d[id=="R_d4o2OrwcVxrjEn7"]$soccer <- "None of the above"

# survey id: R_8c7WtV8AzXTDACl - missing fauci
d[id=="R_8c7WtV8AzXTDACl"]$soccer <- "None of the above"
####################

# After reviewing all the missing data, it seems that the option "None of the above" wasn't registered in any of the option

d[georgians=="" & finished=="True"]$georgians <- "None of the above"
d[energy=="" & finished=="True"]$energy <- "None of the above"
d[soccer=="" & finished=="True"]$soccer <- "None of the above"
d[fauci=="" & finished=="True"]$fauci <- "None of the above"
d[pollution=="" & finished=="True"]$pollution <- "None of the above"
d[election=="" & finished=="True"]$election <- "None of the above"

#duplicate rows
## ONLY FOR TN
## Respondent id-R_2dWlnlkYnG9epqt is a non-complier
non_compliers_tn<-c("R_0IJ2yZ90ItKq2Fb")
d<-d[!(id %in% non_compliers_tn),]

## ONLY FOR FP and FN
non_compliers_fp<-0
paste("We found",non_compliers_fp," non-compliers")



## Check for attrition
# People who received the treatment and then stopped
attrition_count <- length(d[finished=="False" & tweet_submit_time!=""]$id)

# People who received the treatment
total_responses <- length(d[tweet_submit_time!=""]$id)

attrition_proportion=attrition_count/total_responses
paste("We found a ", round(attrition_proportion*100, digits = 2),"% attrition.")

# Remove respondents that dropped (attrition + drop out)
respondent_drop <- which(d[,finished=="False"])
d <- d[-respondent_drop,]

# Save emails of raffle respondents
# write.csv(d$email,"../data/interim/tp_emails.csv", row.names = FALSE)
# write.csv(d$email,"../data/interim/tn_emails.csv", row.names = FALSE)
# write.csv(d$email,"../data/interim/fp_emails.csv", row.names = FALSE)
#write.csv(d$email,"../data/interim/fn_emails.csv", row.names = FALSE)

# Data type conversion
d$start_date <- as.Date(d$start_date, format = "%Y-%m-%d %H:%M:%s")
d$end_date <- as.Date(d$end_date, format = "%Y-%m-%d %H:%M:%s")
d$tweet_submit_time <- as.integer(d$tweet_submit_time)
d$test_submit_time <- as.integer(d$test_submit_time)
```

We can also see a histogram of time looking at the tweets and time responding the test.

```{r}
#par(mfrow=c(1,2))

h1 <- hist(d$tweet_submit_time,
    main = "Time reading Tweets",
    xlab = "seconds",
    breaks = length(d$tweet_submit_time)/5)

h2<- hist(d$test_submit_time,
    main = "Time responding the test",
    xlab = "seconds",
    breaks = length(d$test_submit_time)/5)


```


Data seems reasonably normal. Now let's remove variable that are not of interest.

```{r}
# Keep only columns of interest
colnames(d)
cols_to_remove <- c("start_date","end_date",
                   "finished","finished_date","id","email",
                   "tweet_first_click_time","tweet_last_click_time","tweet_click_count",
                   "test_first_click_time","test_last_click_time","test_click_count")

cols_to_drop_ix<-which(colnames(d) %in% cols_to_remove)

d[,cols_to_drop_ix]<- list(NULL)
```

Let's include the treatment columns to the data.

```{r}
# IMPORTANT Adding factors per the current dataset of a specific treatment
# d$truth<- as.factor("true")
d$truth<- as.factor("false")

# d$sentiment<- as.factor("positive")
#d$sentiment<- as.factor("negative")



# Correct answers for test: FACT
d[ , bin_stimulus := ifelse(stimulus == "Green House Gases", yes = 1, no = 0)]
d[ , bin_georgians := ifelse(georgians == "None of the above", yes = 1, no = 0)]
d[ , bin_energy := ifelse(energy == "Natural Gas", yes = 1, no = 0)]
d[ , bin_soccer := ifelse(soccer == "Unknown", yes = 1, no = 0)]
d[ , bin_fauci := ifelse(fauci == "National Geographic Society", yes = 1, no = 0)]
d[ , bin_pollution := ifelse(pollution == "6th largest", yes = 1, no = 0)]
d[ , bin_election := ifelse(election == "2024", yes = 1, no = 0)]


# # Correct answers for test: FAKE
d[ , bin_stimulus := ifelse(stimulus == "Green House Gases", yes = 1, no = 0)]
d[ , bin_georgians := ifelse(georgians == "70", yes = 1, no = 0)]
d[ , bin_energy := ifelse(energy == "Wind", yes = 1, no = 0)]
d[ , bin_soccer := ifelse(soccer == "Pfizer vaccine", yes = 1, no = 0)]
d[ , bin_fauci := ifelse(fauci == "None of the above", yes = 1, no = 0)]
d[ , bin_pollution := ifelse(pollution == "1st largest", yes = 1, no = 0)]
d[ , bin_election := ifelse(election == "2024", yes = 1, no = 0)]

d$total_correct <- d%>% select(c("bin_stimulus", "bin_georgians","bin_energy","bin_soccer","bin_fauci","bin_pollution", "bin_election")) %>% rowSums()
```

We can see the distribution of correct answers.

```{r}
h3 <- hist(d$total_correct,
   main = "Correct answers distribution (max 7)",
    xlab = "correct answers",
    breaks = length(d$total_correct)/4)
```

Now we can save the file for future reference and stacking.

```{r}
# Save to CSV
write.csv(d,"../data/interim/tp_data.csv", row.names = FALSE)
write.csv(d,"../data/interim/tn_data.csv", row.names = FALSE)
write.csv(d,"../data/interim/fp_data.csv", row.names = FALSE)
write.csv(d,"../data/interim/fn_data.csv", row.names = FALSE)
```

Finally, we can stack all of the individual files.

```{r}
files <- c("tp_data.csv","tn_data.csv","fp_data.csv","fn_data.csv")

df_fn<-data.table::fread("../data/interim/fn_data.csv")
df_fp<-data.table::fread("../data/interim/fp_data.csv")
df_tn<-data.table::fread("../data/interim/tn_data.csv")
df_tp<-data.table::fread("../data/interim/tp_data.csv")

df<-rbind(df_tp,df_tn,df_fp,df_fn)

write.csv(df,"../data/processed/tweets_data.csv", row.names = FALSE)

```

START HERE: Let's confirm data was saved properly

```{r}
df<-data.table::fread("../data/processed/tweets_data.csv")

df$age <- gsub("[a-zA-Z]", "", df$age)
    
  
df$truth_f <- as.factor(df$truth)
df$sentiment_f <- as.factor(df$sentiment)
df$age_f <- as.factor(df$age)
df$education_f <- as.factor(df$education)
df$gender_f <- as.factor(df$gender)
df[ truth == 'TRUE' & sentiment == 'positive', treatment_group := 'TP'] 
df[ truth == 'TRUE' & sentiment == 'negative', treatment_group := 'TN'] 
df[ truth == 'FALSE' & sentiment == 'positive', treatment_group := 'FP'] 
df[ truth == 'FALSE' & sentiment == 'negative', treatment_group := 'FN'] 
df$treatment_group_f <- as.factor(df$treatment_group)

df_tp <- df[ truth == 'TRUE' & sentiment == 'positive', ]
df_tn <- df[ truth == 'TRUE' & sentiment == 'negative', ]
df_fp <- df[ truth == 'FALSE' & sentiment == 'positive', ]
df_fn <- df[ truth == 'FALSE' & sentiment == 'negative', ]

```

```{r}
tail(df)
```


```{r}
par(mfrow=c(2,3))

plot(df$age_f, main='Age - Overall.')
plot(df$education_f, main='Educ. - Overall.')
plot(df$gender_f, main='Gender - Overall.')
plot(as.factor(df$treatment_group), main='Treatment Group.')
hist(df$total_correct, main='Total Corr. - Overall', xlab='Total Correct')
```
```{r}
#par(mfrow=c(1,2))

#qplot(df$total_correct, 
#      geom="histogram",
#      main = "Histogram: Total Correct Count",
#      xlab = "Correct Response Count(max = 7)",
#      ylab = "Occurrence Count",
#      fill = I("grey"),
#      col = I("blue"),
#      xlim=c(NA,8)) 



g1 <- ggplot(df, aes(x=age_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue"))+
  ggtitle("Age") +
  xlab("Age Bucket") + ylab("Count") 

g2 <- ggplot(df, aes(x=education_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("Education Level") +
  xlab("Highest Level") + ylab("Count")
g3 <- ggplot(df, aes(x=gender_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("Gender") +
  xlab("Gender") + ylab("Count")
g4 <- ggplot(df, aes(x=total_correct)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("Total Correct") +
  xlab("Correct Responses") + ylab("Count")
grid.arrange(g1, g2, g3, g4, nrow = 2)

#plot(df$education_f, main='Educ. - Overall.')
#plot(df$gender_f, main='Gender - Overall.')
#plot(as.factor(df$treatment_group), main='Treatment Group.')
#hist(df$total_correct, main='Total Corr. - Overall', xlab='Total Correct')
```
```{r}
g0 <- ggplot(df, aes(x=total_correct)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("Overall - Total Correct") +
  xlab("Correct Responses") + ylab("Count")

g1 <- ggplot(df_tp, aes(x=total_correct)) +
  geom_bar( fill = I("grey"),
      col = I("blue"))+
  ggtitle("True-Positive") +
  xlab(NULL) + ylab("Count")

g2 <- ggplot(df_tn, aes(x=total_correct)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("True-Negative") +
  xlab(NULL) + ylab(NULL)

g3 <- ggplot(df_fp, aes(x=total_correct)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Positive") +
  xlab("Correct Responses") + ylab("Count")

g4 <- ggplot(df_fn, aes(x=total_correct)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Negative") +
  xlab("Correct Responses") + ylab(NULL)

grid.arrange(g0, g1, g2, g3, g4, widths = c(2.5, 1.5, 1.5),
  layout_matrix = rbind(c(1, 2, 3), 
                        c(1, 4, 5)))

```

```{r}
a0 <- ggplot(df, aes(x=age_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("Overall - Age Bucket") +
  xlab("Age Bucket") + ylab("Count") +
  theme(axis.text.x = element_text(angle=90))
a1 <- ggplot(df_tp, aes(x=age_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue"))+
  ggtitle("True-Positive") +
  xlab(NULL) + ylab("Count") +
  theme(axis.text.x = element_text(angle=90))
a2 <- ggplot(df_tn, aes(x=age_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("True-Negative") +
  xlab(NULL) + ylab(NULL) +
  theme(axis.text.x = element_text(angle=90))
a3 <- ggplot(df_fp, aes(x=age_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Positive") +
  xlab("Age Bucket") + ylab("Count")+
  theme(axis.text.x = element_text(angle=90))
a4 <- ggplot(df_fn, aes(x=age_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Negative") +
  xlab("Age Bucket") + ylab(NULL)+
  theme(axis.text.x = element_text(angle=90))

grid.arrange(a0, a1, a2, a3, a4, widths = c(2, 1, 1),
  layout_matrix = rbind(c(1, 2, 3), 
                        c(1, 4, 5))) 

```
```{r}
gdr0 <- ggplot(df, aes(x=gender_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("Overall - Gender") +
  xlab("Gender") + ylab("Count")
gdr1 <- ggplot(df_tp, aes(x=gender_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue"))+
  ggtitle("True-Positive") +
  xlab(NULL) + ylab("Count")
gdr2 <- ggplot(df_tn, aes(x=gender_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("True-Negative") +
  xlab(NULL) + ylab(NULL)
gdr3 <- ggplot(df_fp, aes(x=gender_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Positive") +
  xlab("Gender") + ylab("Count")
gdr4 <- ggplot(df_fn, aes(x=gender_f)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Negative") +
  xlab("Gender") + ylab(NULL)

grid.arrange(gdr0, gdr1, gdr2, gdr3, gdr4, widths = c(2, 1, 1),
  layout_matrix = rbind(c(1, 2, 3), 
                        c(1, 4, 5)))
```
```{r}
gdr0 <- ggplot(df, aes(x=education)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("Overall - Education") +
  xlab("Education") + ylab("Count")+
  theme(axis.text.x = element_text(angle=90))
gdr1 <- ggplot(df_tp, aes(x=education)) +
  geom_bar( fill = I("grey"),
      col = I("blue"))+
  ggtitle("True-Positive") +
  xlab(NULL) + ylab("Count")+
  theme(axis.text.x = element_text(angle=90))
gdr2 <- ggplot(df_tn, aes(x=education)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("True-Negative") +
  xlab(NULL) + ylab(NULL)+
  theme(axis.text.x = element_text(angle=90))
gdr3 <- ggplot(df_fp, aes(x=education)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Positive") +
  xlab("Education") + ylab("Count")+
  theme(axis.text.x = element_text(angle=90))
gdr4 <- ggplot(df_fn, aes(x=education)) +
  geom_bar( fill = I("grey"),
      col = I("blue")) +
  ggtitle("False-Negative") +
  xlab("Education") + ylab(NULL)+
  theme(axis.text.x = element_text(angle=90))
grid.arrange(gdr0, gdr1, gdr2, gdr3, gdr4, widths = c(2, 1, 1),
  layout_matrix = rbind(c(1, 2, 3), 
                        c(1, 4, 5)))
```

